Oregon 2013 Medication for Asthma Inhalers

### 2017-07-07 r_scripts/analysis/oregon_medication.Rmd
doing distributed lag
update revised zipcode and DLM

```{r library, include=FALSE, echo=FALSE}
library(dplyr)
library(data.table)
library(readxl) 
library(readr)
library(survival)
library(ggplot2)
library(htmlTable)
library(knitr)
library(splines)
```

@@

## Import Asthma NDC table
read_path2 <- paste0("../../../data/data_new/ndc_2013_beta_2_agonists.csv")
asthma_ndc <- read_csv(read_path2, col_names = FALSE)

colnames(asthma_ndc) <- c("ndc_code", "brand_name", "generic_product_name", 
                          "route", "category", "drug_id")

# convert to vector
ndc_codes <- as.vector(as.matrix(asthma_ndc$ndc_code))

@@

```{r time stratified, eval=FALSE, echo=FALSE}
### Time stratified ------------------------------------------------------------
# read in zipcode level populatoin-weighted pm
read_path5 <- paste0('C:/Users/jyliu/Desktop/local_git_repo/oregon_wildfire_new/data/Oregon_PM/zip_pm_to_merge_with_acap.csv')

@@

write_path4 <- paste0("../data_new/medication/oregon_ndc_time_strat_casecrossover.csv")
write_csv(outcome_casecross, write_path4)

@@

```{r descriptive table, echo = F, message = F, warning = F, results='asis'}
read_path6 <- paste0("../data_new/medication/oregon_ndc_time_strat_casecrossover.csv")
asthma_ndc <- read_csv(read_path6)

@@

# dataframe to loop through
asthma_ndc_df <- data.frame(asthma_ndc)

@@

# dataframe for analysis creation
# bind columns back together 
df_analysis <- asthma_ndc_df %>%
  filter(!is.na(wrf_smk_pm_zip)) %>% 
  # only look at outcomes
  filter(outcome == 1) %>%
  # add another row that makes sure there is a person <15 in the dataframe
  # tricking xtabs to produce a 0 cell for the outcome for age <15
  add_row(outcome = 0, age_ind = 0)

@@

# extract covariates from dataframe
covariates_df <- asthma_ndc_df[, c(1:26, 74:84)]

# extract pm values and divide by 10 and ordered
#which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
pm_estimates_df <- asthma_ndc_df[, c(87, 92, 91, 93)]/10  # create 10 unit increases

@@

### Overall pharmacy visit time series

The time series for the number of pharmacy in each day and separated with different counties. For convenience and representativeness, I choose some of the largest counties to show the visit features.

The overall time series plot shows a time series plot.

```{r time series, warning=FALSE, message=FALSE, echo=FALSE}
read_path7 <- paste0("../data_new/medication/oregon_asthma_ndc.csv")
or_ndc_claim <- read_csv(read_path7)

or_ndc_asthma <- or_ndc_claim %>%
  select(personkey, num_visit, clmid, line, ZIP, dates)

### Join county with zip
read_path <- paste0('../data_new/update/or_zip_county_prop.csv') 
or_zip_county <- read_csv(read_path)

or_zip_county <- or_zip_county %>%
  select(zip, county_name) %>%
  rename(county = county_name) %>%
  rename(ZIP = zip) %>%
  full_join(oregon_fips, by = "county") %>%
  select(ZIP, county, state, st_county_fips) %>%
  rename(fips = st_county_fips)

or_zip_county$county[which(or_zip_county$county=="Hood.River")] <- "Hood River"

# Import population weighted county PM2.5
pm_path <- paste0("../data_new/county_data/or_county_pop_wt_pm.csv")
county_pm <- read_csv(pm_path)

# create a working dataset 'or_2013'
or_ndc_county_time_series <- or_ndc_asthma %>% 
  # join in the washington county name based on CHARS county identifier
  left_join(or_zip_county, by = "ZIP") %>% 
  # rename 'admit_date_impute' to 'date'
  rename(date = dates) %>%
  # rename(date = admit_date_impute) %>% 
  # group by county of residence and date
  group_by(county, date) %>% 
  # sum up each primary diagnosis for each outcome for each day for each county
  summarise(n_obs = n()) %>% 
  # join daily county estimates of pm2.5
  left_join(county_pm, by = c('date', 'county'))

# which(is.na(or_ndc_county_time_series$county))
# or_ndc_county_time_series[113,]
# which(or_zip_county$ZIP=="97078")

# summary(or_ndc_asthma)
# summary(or_ndc_county_time_series)
# summary(as.factor(or_ndc_county_time_series$county)) # 125 is missing out of 3592 (3.16%)

# remove NA values
# which(is.na(or_ndc_county_time_series$county)) # 3468-3592
or_ndc_county_time_series <- or_ndc_county_time_series[-c(3468:3592),]

# write permanent dataframe

write_path5 <- paste0("../data_new/medication/or_ndc_county_time_series.csv")
write_csv(or_ndc_county_time_series, write_path5)

## Check for missing values in sparsely populated counties. set to 0 (But no)
# check_missing <- filter(or_ndc_county_time_series, is.na(n_obs))
# mutate each to fill in missing values (Still no)
# check2 <- check_missing %>% mutate_each(funs(new = ifelse(is.na(.), 0, .)), 
#                                         n_obs:wrf_temp)

plot_ts <- ggplot(or_ndc_county_time_series, aes(y =n_obs, x = date)) +
           geom_jitter()

print(plot_ts)
# save figure
ggsave("../plot_new/ndc_ts.pdf", plot = plot_ts, 
       width = 12, height = 8, units = "in")
```

### County time series

The dotted time series plot of pharmacy visits by counties shows a large and distinct strata dotted time series in some big counties (such as Multnomah), I guess it may be the difference of visits between weekday and weekend. The lined time series plot also shows fluctuation. So we choose some larger counties (with more visits) to research on the effect of weekend and weekay.

```{r, warning=FALSE, message=FALSE,echo=FALSE}

plot_county <- ggplot(or_ndc_county_time_series, aes(y =n_obs, x = date)) +
               geom_jitter() + facet_wrap(~county)

plot_county_line <- ggplot(or_ndc_county_time_series, aes(y =n_obs, x = date)) +
                    facet_wrap(~county) + geom_line()

print(plot_county)
print(plot_county_line)
# save figure
ggsave("../plot_new/ndc_county_ts.pdf", plot = plot_county, 
       width = 12, height = 8, units = "in")
ggsave("../plot_new/ndc_county_ts_line.pdf", plot = plot_county_line, 
       width = 12, height = 8, units = "in")

@@

### Time series of weekday and weekend in larger counties.
We check the time series for Multnomah, and it indeed shows more visits of weekdays than that in weekends (except some holidays such like Labor Day, Independent Day, etc.) We may choose the new covariates for weekday and weekend.

```{r echo = FALSE, message=FALSE, warning=FALSE}
or_ndc_county_time_series_week <- or_ndc_county_time_series%>%
  mutate(day_admit = as.factor(weekdays(date)),
         month_admit = month(date)) %>%
  mutate(outcome = 1)

# Also check for Multnomah
multnomah_df <- or_ndc_county_time_series_week %>%
  filter(county == "Multnomah")

plot_mult_week <- ggplot(multnomah_df, aes(y =n_obs, x = date)) +
           geom_jitter() + facet_wrap(~day_admit)

print(plot_mult_week)

ggsave("../plot_new/multnomah_ts_week.pdf", plot = plot_mult_week, 
       width = 12, height = 8, units = "in")
```

### Time series of smoke exposure
The plot shows the times series of smoke PM 2.5 exposure. It seems that at the beginning of August or at the end of July the smoke has a peak. Also during the August the smoke is very high. The green one shows the WRF-Chem method; The bule one shows the Krigging method; And the red one, which we focus on, shows the GWR method.

```{r echo = FALSE, message=FALSE, warning=FALSE}

plot_smoke_ts_wrf <- ggplot(or_ndc_county_time_series_week, aes(y =wrf_smk_pm, x = date)) +
           geom_jitter(color="red")
# print(plot_smoke_ts_wrf)

plot_smoke_ts_krig <- ggplot(or_ndc_county_time_series_week, aes(y =krig_smk_pm, x = date)) +
           geom_jitter() 
# print(plot_smoke_ts_krig)

plot_smoke_ts_gwr <- ggplot(or_ndc_county_time_series_week, aes(y =geo_smk_pm, x = date)) +
           geom_jitter() 
# print(plot_smoke_ts_gwr)

plot_smk <- ggplot(or_ndc_county_time_series_week, aes(x = date)) +
  geom_jitter(aes(y =wrf_smk_pm, x = date), color='green', size = 1) + 
  geom_jitter(aes(y =krig_smk_pm, x = date), color='blue', size = 1) + 
  geom_jitter(aes(y =geo_smk_pm, x = date), color="red", size = 1)  +
  ylab("Smoke PM 2.5")

print(plot_smk)

@@

```{r DLM, echo = FALSE, message=FALSE, warning=FALSE}

which(colnames(asthma_ndc_df)=="geo_wt_pm_zip") # 88
which(colnames(asthma_ndc_df)=="geo_wt_pm_lag1_zip") #115

asthma_ndc_lag <- asthma_ndc_df %>%
  filter(gender!="U") %>%
  filter(outcome==1) %>%
  select(88, 115:121) # GWR-method

lag_max <- 7
n <- length(asthma_ndc_lag[,1])

asthma_ndc_lag <- as.matrix(asthma_ndc_lag)

# true DL function
DLfunc <- exp(-(0:lag_max)/2)

# simulate outcomes
y <- drop(asthma_ndc_lag%*%DLfunc) # + rnorm(n)

B <- ns(0:lag_max, df =4, intercept=TRUE)

# the nrow B should equal ncol(pm)
dim(B)
dim(asthma_ndc_lag)

asthma_ndc_lag_B <- asthma_ndc_lag%*%B
fit <- lm(y~asthma_ndc_lag_B)
coef(fit)

dlparms <- grep("asthma_ndc_lag_B",names(coef(fit)))
DLestimate<- data.frame(estimate=B%*%coef(fit)[dlparms])
DLvar <- B%*%vcov(fit)[dlparms,dlparms]%*%t(B)
DLestimate$SE <- sqrt(diag(DLvar))
DLestimate$lower <- DLestimate$estimate + DLestimate$SE * qt(1-0.975, df=fit$df.residual)
DLestimate$upper <- DLestimate$estimate + DLestimate$SE * qt(0.975, df=fit$df.residual)
DLestimate

DLestimate$lag <- 0:lag_max
DLestimate$truth <- DLfunc
p <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper))
p <- p + geom_ribbon(alpha=.5) + geom_line(size=2)
p <- p + geom_line(aes(x=lag, y=truth), linetype=2, size=2)
p <- p + theme_bw()
p

cumulative <- sum(DLestimate$estimate)
cumulative
## [1] 2.485711
cumulative_se <- sqrt(sum(DLvar))
cumulative_se
## [1] 0.0001887442
cumulative_CI <- cumulative + cumulative_se * qt(c(1-0.975,0.975), df=fit$df.residual)
cumulative_CI
## [1] 2.485341 2.486080

>>>>

```{r library, include=FALSE, echo=FALSE}
library(tidyverse)
# library(data.table)
library(survival)
library(ggplot2)
library(htmlTable)
library(knitr)
library(broom)
library(lme4)
library(lubridate) # date

# two-stage distributed lag model libraries
library(dlnm)
library(mvmeta)
library(splines)

@@

## Import Asthma NDC table
read_path2 <- paste0("../../../data/data_new/ndc_2013_beta_2_agonists.csv")
inhaler_ndc <- read_csv(read_path2, col_names = FALSE)

colnames(inhaler_ndc) <- c("ndc_code", "brand_name", "generic_product_name", 
                          "route", "category", "drug_id")

# convert to vector
ndc_codes <- as.vector(as.matrix(inhaler_ndc$ndc_code))

@@

```{r time stratified zip for casecrossover, eval=FALSE, echo=FALSE}
### Time stratified ------------------------------------------------------------
# read in zipcode level populatoin-weighted pm
read_path5 <- paste0('C:/Users/jyliu/Desktop/local_git_repo/oregon_wildfire_new/data/Oregon_PM/zip_pm_to_merge_with_acap.csv')

@@

write_path4 <- paste0("../data_new/medication/oregon_ndc_time_strat_casecrossover_zip.csv")
write_csv(outcome_casecross, write_path4)

## Time stratified for time series (Population-Weighted PM~2.5~ Assigned by County)

```{r time stratified county for ts, echo=FALSE, include=FALSE, eval=FALSE}
read_path7 <- paste0("../data_new/medication/oregon_asthma_ndc.csv")
or_ndc_claim <- read_csv(read_path7)

or_ndc_inhaler <- or_ndc_claim %>%
  select(personkey, num_visit, clmid, line, ZIP, dates) %>%
  rename(ZIPCODE = ZIP)

### Join county with zip
# read_path <- paste0('../data_new/update/or_zip_county_prop.csv') 
# or_zip_county <- read_csv(read_path)

read_path <- paste0('../instructions/oregon_zip_county.csv')
or_zip_county <- read_csv(read_path)

# or_zip_county$county_name[which(or_zip_county$county_name=="Hood.River")] <- "Hood River"

or_zip_county <- or_zip_county %>%
  select(zip, county_name) %>%
  rename(county = county_name) %>%
  rename(ZIPCODE = zip) %>%
  full_join(oregon_fips, by = "county") %>%
  select(ZIPCODE, county, state, st_county_fips) %>%
  rename(fips = st_county_fips) # 484

county_new <- c("Washington", "Douglas", "Lane", "Deschutes")

zip_new <- or_ndc_inhaler %>%
  filter(!(ZIPCODE %in% or_zip_county$ZIPCODE)) %>%
  select(ZIPCODE) %>%
  unique() %>%
  arrange(ZIPCODE) %>%
  mutate(county = county_new)

# summary(as.factor(zip_new$ZIPCODE))
# 97003 97471 97475 97703 
#    48   138     5    16

or_zip_county_new <- or_zip_county %>%
  select(ZIPCODE, county) %>%
  bind_rows(zip_new)

write_path <- paste0("../data_new/county_data/zip_to_county_new.csv")
write_csv(or_zip_county_new, write_path)

read_path9 <- paste0("../data_new/county_data/or_census_pop_est_county.csv")
or_pop <- read_csv(read_path9)

or_pop <- or_pop %>%
  rename(county = GEO_display_label, pop = respop72013) %>%
  select(county, pop) 
or_pop$county <- gsub(" County, Oregon", "", as.character(factor(or_pop$county)))

### import county data
read_path_county_pm <- paste0('C:/Users/jyliu/Desktop/local_git_repo/oregon_wildfire_new/data_new/county_data/or_county_pop_wt_pm.csv')

county_smoke <- read_csv(read_path_county_pm) # 63801 rows

# descriptives of the two smoke datasets
summary(county_smoke)

county_smoke$county[which(county_smoke$county=="Hood.River")] <- "Hood River"
county_smoke_w_lag <- county_smoke %>% arrange(county, date) %>%
  group_by(county) %>% 
  mutate(wrf_f_pm_lag1 = lag(wrf_pm, 1, order_by = county), 
         wrf_f_pm_lag2 = lag(wrf_pm, 2, order_by = county),
         wrf_f_pm_lag3 = lag(wrf_pm, 3, order_by = county),
         wrf_f_pm_lag4 = lag(wrf_pm, 4, order_by = county),
         wrf_f_pm_lag5 = lag(wrf_pm, 5, order_by = county),
         wrf_f_pm_lag6 = lag(wrf_pm, 6, order_by = county),
         wrf_f_pm_lag7 = lag(wrf_pm, 7, order_by = county),
         wrf_nf_pm_lag1 = lag(wrf_nf_pm, 1, order_by = county),
         wrf_nf_pm_lag2 = lag(wrf_nf_pm, 2, order_by = county),
         wrf_nf_pm_lag3 = lag(wrf_nf_pm, 3, order_by = county),
         wrf_nf_pm_lag4 = lag(wrf_nf_pm, 4, order_by = county),
         wrf_nf_pm_lag5 = lag(wrf_nf_pm, 5, order_by = county),
         wrf_nf_pm_lag6 = lag(wrf_nf_pm, 6, order_by = county),
         wrf_nf_pm_lag7 = lag(wrf_nf_pm, 7, order_by = county),
         wrf_smk_pm_lag1 = lag(wrf_smk_pm, 1, order_by = county),
         wrf_smk_pm_lag2 = lag(wrf_smk_pm, 2, order_by = county),
         wrf_smk_pm_lag3 = lag(wrf_smk_pm, 3, order_by = county),
         wrf_smk_pm_lag4 = lag(wrf_smk_pm, 4, order_by = county),
         wrf_smk_pm_lag5 = lag(wrf_smk_pm, 5, order_by = county),
         wrf_smk_pm_lag6 = lag(wrf_smk_pm, 6, order_by = county),
         wrf_smk_pm_lag7 = lag(wrf_smk_pm, 7, order_by = county),
         geo_wt_pm_lag1 = lag(geo_wt_pm, 1, order_by = county),
         geo_wt_pm_lag2 = lag(geo_wt_pm, 2, order_by = county),
         geo_wt_pm_lag3 = lag(geo_wt_pm, 3, order_by = county),
         geo_wt_pm_lag4 = lag(geo_wt_pm, 4, order_by = county),
         geo_wt_pm_lag5 = lag(geo_wt_pm, 5, order_by = county),
         geo_wt_pm_lag6 = lag(geo_wt_pm, 6, order_by = county),
         geo_wt_pm_lag7 = lag(geo_wt_pm, 7, order_by = county),
         krig_pm_lag1 = lag(krig_pm, 1, order_by = county),
         krig_pm_lag2 = lag(krig_pm, 2, order_by = county),
         krig_pm_lag3 = lag(krig_pm, 3, order_by = county),
         krig_pm_lag4 = lag(krig_pm, 4, order_by = county),
         krig_pm_lag5 = lag(krig_pm, 5, order_by = county), 
         krig_pm_lag6 = lag(krig_pm, 6, order_by = county),
         krig_pm_lag7 = lag(krig_pm, 7, order_by = county), 
         background_pm_lag1 = lag(background_pm, 1, order_by = county),
         background_pm_lag2 = lag(background_pm, 2, order_by = county),
         background_pm_lag3 = lag(background_pm, 3, order_by = county),
         background_pm_lag4 = lag(background_pm, 4, order_by = county),
         background_pm_lag5 = lag(background_pm, 5, order_by = county), 
         background_pm_lag6 = lag(background_pm, 6, order_by = county),
         background_pm_lag7 = lag(background_pm, 7, order_by = county), 
         geo_smk_pm_lag1 = lag(geo_smk_pm, 1, order_by = county),
         geo_smk_pm_lag2 = lag(geo_smk_pm, 2, order_by = county),
         geo_smk_pm_lag3 = lag(geo_smk_pm, 3, order_by = county),
         geo_smk_pm_lag4 = lag(geo_smk_pm, 4, order_by = county),
         geo_smk_pm_lag5 = lag(geo_smk_pm, 5, order_by = county),
         geo_smk_pm_lag6 = lag(geo_smk_pm, 6, order_by = county),
         geo_smk_pm_lag7 = lag(geo_smk_pm, 7, order_by = county),
         krig_smk_pm_lag1 = lag(krig_smk_pm, 1, order_by = county),
         krig_smk_pm_lag2 = lag(krig_smk_pm, 2, order_by = county),
         krig_smk_pm_lag3 = lag(krig_smk_pm, 3, order_by = county),
         krig_smk_pm_lag4 = lag(krig_smk_pm, 4, order_by = county),
         krig_smk_pm_lag5 = lag(krig_smk_pm, 5, order_by = county),
         krig_smk_pm_lag6 = lag(krig_smk_pm, 6, order_by = county),
         krig_smk_pm_lag7 = lag(krig_smk_pm, 7, order_by = county),
         wrf_temp_lag1 = lag(wrf_temp, 1, order_by = county),
         wrf_temp_lag2 = lag(wrf_temp, 2, order_by = county),
         wrf_temp_lag3 = lag(wrf_temp, 3, order_by = county),
         wrf_temp_lag4 = lag(wrf_temp, 4, order_by = county),
         wrf_temp_lag5 = lag(wrf_temp, 5, order_by = county), 
         wrf_temp_lag6 = lag(wrf_temp, 6, order_by = county),
         wrf_temp_lag7 = lag(wrf_temp, 7, order_by = county)) %>% 
  ungroup(county) %>% 
  setNames(paste(colnames(.), "county", sep="_"))  %>% 
  rename(county = county_county, date = date_county, fips = fips_county) %>%
  select(-1)

or_ndc_county_ts <- or_ndc_inhaler %>%
  rename(date = dates) %>%
  mutate(day = ifelse(weekdays(date) == "Monday" |
                      weekdays(date) == "Tuesday" |
                      weekdays(date) == "Wednesday"|
                      weekdays(date) == "Thursday"|
                      weekdays(date) == "Friday", "weekday",
               ifelse(weekdays(date) == "Saturday" |
                      weekdays(date) == "Sunday", "weekend", NA))) %>%
  # join with county-level population
  left_join(or_zip_county_new, by = "ZIPCODE") %>% # 321 missing 
  filter(!is.na(county)) %>% # no missing value
  left_join(or_pop, by = "county") %>%
  left_join(county_smoke_w_lag, by = c("date", "county")) %>%
  arrange(personkey, date) %>% # order by id and fromdate
  group_by(county, date) %>% 
  # sum up each primary diagnosis for each outcome for each day for each county
  summarise(n_obs = n()) %>% 
  # join daily county estimates of pm2.5
  left_join(county_smoke_w_lag, by = c('date', 'county'))

# summary(as.factor(or_ndc_county_ts$county))

or_ndc_county_ts <- or_ndc_county_ts %>%
  left_join(or_pop, by = "county") %>%
  mutate(geo_smk_pm10 = geo_smk_pm_county/10) %>%
  mutate(rate_per_100k = (n_obs/pop)*100000) %>%
  mutate(day = ifelse(weekdays(date) == "Monday" |
                      weekdays(date) == "Tuesday" |
                      weekdays(date) == "Wednesday"|
                      weekdays(date) == "Thursday"|
                      weekdays(date) == "Friday", "weekday",
               ifelse(weekdays(date) == "Saturday" |
                      weekdays(date) == "Sunday", "weekend", NA)))

or_ndc_county_ts$day[which(or_ndc_county_ts$date=="2013-05-27")] <- "weekend"
or_ndc_county_ts$day[which(or_ndc_county_ts$date=="2013-07-04")] <- "weekend"
or_ndc_county_ts$day[which(or_ndc_county_ts$date=="2013-09-02")] <- "weekend"

write_path5 <- paste0("../data_new/medication/or_ndc_county_time_series.csv")
write_csv(or_ndc_county_ts, write_path5)

@@

```{r descriptive table, echo = F, message = F, warning = F, results='asis'}
read_path6 <- paste0("../data_new/medication/oregon_ndc_time_strat_casecrossover_zip.csv")
inhaler_ndc <- read_csv(read_path6)
# asthma_ndc <- fread(read_path6, showProgress = T) # faster, but seems error in overall plot analysis

@@

# dataframe to loop through
inhaler_ndc_df <- data.frame(inhaler_ndc)

@@

# dataframe for analysis creation
# bind columns back together 
df_analysis <- inhaler_ndc_df %>%
  filter(!is.na(wrf_smk_pm_zip)) %>% 
  # only look at outcomes
  filter(outcome == 1) %>%
  # add another row that makes sure there is a person <15 in the dataframe
  # tricking xtabs to produce a 0 cell for the outcome for age <15
  add_row(outcome = 0, age_ind = 0)

@@

# extract covariates from dataframe
covariates_df <- inhaler_ndc_df[, c(1:26, 74:84)]

# extract pm values and divide by 10 and ordered
#which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
pm_estimates_df <- inhaler_ndc_df[, c(87, 92, 91, 93)]/10  # create 10 unit increases

@@

### County Time series plots of rate

### Prepare (already done in previous time stratified work)

1. Make 10 units' geo smoke variable
2. Make county population variable (to calculate rate)
3. Make weekend/weekday variables

```{r time series, warning=FALSE, message=FALSE, echo=FALSE}
read_path5 <- paste0("../data_new/medication/or_ndc_county_time_series.csv")
or_ndc_county_ts <- read_csv(read_path5)

plot_ts <- ggplot(or_ndc_county_ts, aes(y =rate_per_100k, x = date)) +
           geom_jitter()

print(plot_ts)

# save figure
ggsave("../plot_new/ndc_rate_ts.pdf", plot = plot_ts, 
       width = 12, height = 8, units = "in")

plot_county <- ggplot(or_ndc_county_ts, aes(y =rate_per_100k, x = date)) +
               geom_jitter() + facet_wrap(~county)

plot_county_line <- ggplot(or_ndc_county_ts, aes(y =rate_per_100k, x = date)) +
                    facet_wrap(~county) + geom_line()

print(plot_county)
print(plot_county_line)
# save figure
ggsave("../plot_new/ndc_county_rate_ts.pdf", plot = plot_county, 
       width = 12, height = 8, units = "in")
ggsave("../plot_new/ndc_county_rate_ts_line.pdf", plot = plot_county_line, 
       width = 12, height = 8, units = "in")

### check if the time series has the association with get smoke
a <- glm(rate_per_100k ~ geo_smk_pm10, data = or_ndc_county_ts, family = "poisson")
summary(a)

plot_health_smk <- ggplot(or_ndc_county_ts, aes(x = geo_smk_pm10, y = log(rate_per_100k+0.1))) +
  geom_point(aes(colour = c("red")), alpha = 0.2,  size = 1) + 
  # scale_colour_manual(values = c("red", "blue")) +
  # truncated y axis
  #scale_y_continuous(limits = c(0, 1.5)) +
  # facet_wrap(~county, scale  = "free") +
  ylab("Log Outcome Rate per 100,000 Persons") +
  xlab("GWR Smoke PM2.5 per 10*10ug/m^3") +
  theme_bw()

# unnecessary
plot_county_health_smk <- ggplot(or_ndc_county_ts, aes(x = geo_smk_pm10, y = log(rate_per_100k+0.1))) +
  geom_point(aes(colour = county), alpha = 0.2,  size = 1) + 
  # scale_colour_manual(values = c("red", "blue")) +
  # truncated y axis
  #scale_y_continuous(limits = c(0, 1.5)) +
  facet_wrap(~county, scale  = "free") +
  ylab("Log Outcome Rate per 100,000 Persons") +
  xlab("GWR Smoke PM2.5 per 10*10ug/m^3") +
  theme_bw()

plot(plot_health_smk)
plot(plot_county_health_smk)

ggsave("../plot_new/ndc_rate_health_smk.pdf", plot = plot_health_smk, 
       width = 12, height = 8, units = "in")

ggsave("../plot_new/ndc_county_rate_health_smk.pdf", plot = plot_county_health_smk, 
       width = 12, height = 8, units = "in")

@@

### Standard Regression Model

Same-day estimates of association between smoke via GWR and outcomes. County treated as a fixed effect covariate.

1. Quasipoisson

```{r regression model q, message=FALSE, echo=FALSE}
# create empty dataframe
rr_df <- data.frame(matrix(nrow = 1, ncol = 4))
colnames(rr_df) <- c("outcome", "rel_risk", "lower_95", "upper_95")

# vector of outcomes
outcome_list <- "n_obs"

# names 
outcome_names <- c('Inhalers')

# populate outcome column
rr_df[,1] <- factor(outcome_names, levels=unique(outcome_names))

smk_est <- tidy(glm(as.formula(paste0(outcome_list, 
    "~ geo_smk_pm10 + wrf_temp_county + county + 
    day + offset(log(pop))")), 
    or_ndc_county_ts, family="quasipoisson"))[2, ]

rr <- c(round(exp(smk_est[1,2]),5),
        round(exp(smk_est[1,2]-1.96*smk_est[1,3]),5),
        round(exp(smk_est[1,2]+1.96*smk_est[1,3]),5))

rr_df[1,2:4] <- rr

# tried apply function and it might have been easier to just use a loop
# as I needed a for loop to fill the list in to the dataframe
# code is more compact though
# I should check speed (or make a custom function to do this one day)

rr_df

plot_rr <- ggplot(rr_df, aes(x=outcome, y = rel_risk)) +
  geom_point() +
  geom_errorbar(aes(ymin=lower_95, ymax=upper_95), width = 0.2) +
  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
  theme(
   panel.background = element_rect(fill = 'white', colour = 'black'),
   panel.grid.major = element_blank(),
   panel.grid.minor = element_blank(),
   axis.text.x = element_text(angle = 90, vjust = .75)) +
  ylab("Relative Risk") +
  xlab("Outcome") +
  ggtitle("Relative Risk for a 10ug/m^3 increase PM2.5 estimated via GWR Smoke")

print(plot_rr)

ggsave("../plot_new/ndc_rr.pdf", plot = plot_rr, 
       width = 12, height = 8, units = "in")

2. Poisson

```{r regression model,  message=FALSE, echo=FALSE}
# create empty dataframe
rr_df <- data.frame(matrix(nrow = 1, ncol = 4))
colnames(rr_df) <- c("outcome", "rel_risk", "lower_95", "upper_95")

# vector of outcomes
outcome_list <- "n_obs"

# names 
outcome_names <- c('Inhalers')

# populate outcome column
rr_df[,1] <- factor(outcome_names, levels=unique(outcome_names))

smk_est <- tidy(glm(as.formula(paste0(outcome_list, 
    "~ geo_smk_pm10 + wrf_temp_county + county + 
    day + offset(log(pop))")), 
    or_ndc_county_ts, family="poisson"))[2, ]

rr <- c(round(exp(smk_est[1,2]),5),
        round(exp(smk_est[1,2]-1.96*smk_est[1,3]),5),
        round(exp(smk_est[1,2]+1.96*smk_est[1,3]),5))

rr_df[1,2:4] <- rr

rr_df

# tried apply function and it might have been easier to just use a loop
# as I needed a for loop to fill the list in to the dataframe
# code is more compact though
# I should check speed (or make a custom function to do this one day)

plot_rr <- ggplot(rr_df, aes(x=outcome, y = rel_risk)) +
  geom_point() +
  geom_errorbar(aes(ymin=lower_95, ymax=upper_95), width = 0.2) +
  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
  theme(
   panel.background = element_rect(fill = 'white', colour = 'black'),
   panel.grid.major = element_blank(),
   panel.grid.minor = element_blank(),
   axis.text.x = element_text(angle = 90, vjust = .75)) +
  ylab("Relative Risk") +
  xlab("Outcome") +
  ggtitle("Relative Risk for a 10ug/m^3 increase PM2.5 estimated via GWR Smoke")

print(plot_rr)

ggsave("../plot_new/ndc_rr_poi.pdf", plot = plot_rr, 
       width = 12, height = 8, units = "in")

### Mixed models

1. Poisson

```{r mixed model,  message=FALSE, echo=FALSE}
# create empty dataframe
rr_mix_df <- data.frame(matrix(nrow = 1, ncol = 4))
colnames(rr_mix_df) <- c("outcome", "rel_risk", "lower_95", "upper_95")

# vector of outcomes
outcome_list <- "n_obs"

# names 
outcome_names <- c('Inhalers')

# populate outcome column
rr_mix_df[,1] <- factor(outcome_names, levels=unique(outcome_names))

smk_est <- tidy(glmer(as.formula(paste0(outcome_list, 
    "~ geo_smk_pm10 + wrf_temp_county + (1|county) + 
    day + offset(log(pop))")), 
    or_ndc_county_ts, family="poisson"))[2, ]

rr <- c(round(exp(smk_est[1,2]),5),
        round(exp(smk_est[1,2]-1.96*smk_est[1,3]),5),
        round(exp(smk_est[1,2]+1.96*smk_est[1,3]),5))

rr_mix_df[1,2:4] <- rr

rr_mix_df

plot_mix_rr <- ggplot(rr_mix_df, aes(x=outcome, y = rel_risk)) +
  geom_point() +
  geom_errorbar(aes(ymin=lower_95, ymax=upper_95), width = 0.2) +
  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
  theme(
   panel.background = element_rect(fill = 'white', colour = 'black'),
   panel.grid.major = element_blank(),
   panel.grid.minor = element_blank(),
   axis.text.x = element_text(angle = 90, vjust = .75)) +
  ylab("Relative Risk") +
  xlab("Outcome") +
  ggtitle("Relative Risk for a 10ug/m^3 increase PM2.5 estimated via GWR Smoke")

print(plot_rr)

ggsave("../plot_new/ndc_mix_rr_poi.pdf", plot = plot_mix_rr, 
       width = 12, height = 8, units = "in")

@@

### Lag 1

Remove the missing values for GWR smoke. 
But I use GWR smoke of zip for the case crossover study and add the county variable.

```{r DLM, echo = FALSE, message=FALSE, warning=FALSE}
read_path8 <- paste0("../data_new/county_data/zip_to_county_new.csv")
or_zip_to_county <- read_csv(read_path8)

read_path9 <- paste0("../data_new/county_data/or_census_pop_est_county.csv")
or_pop <- read_csv(read_path9)

or_pop <- or_pop %>%
  rename(county = GEO_display_label, pop = respop72013) %>%
  select(county, pop) 
or_pop$county <- gsub(" County, Oregon", "", as.character(factor(or_pop$county)))

inhaler_ndc_lag_df <- inhaler_ndc_df %>%
  left_join(or_zip_to_county, by = "ZIPCODE") %>%
  left_join(or_pop, by = "county") %>%
  mutate(day = ifelse(weekdays(date) == "Monday" |
                      weekdays(date) == "Tuesday" |
                      weekdays(date) == "Wednesday"|
                      weekdays(date) == "Thursday"|
                      weekdays(date) == "Friday", "weekday",
               ifelse(weekdays(date) == "Saturday" |
                      weekdays(date) == "Sunday", "weekend", NA))) %>%
  mutate(geo_smk_pm10_zip = geo_smk_pm_zip/10,
         geo_smk_pm10_lag1_zip =geo_smk_pm_lag1_zip/10,
         geo_smk_pm10_lag2_zip =geo_smk_pm_lag2_zip/10,
         geo_smk_pm10_lag3_zip =geo_smk_pm_lag3_zip/10,
         geo_smk_pm10_lag4_zip =geo_smk_pm_lag4_zip/10,
         geo_smk_pm10_lag5_zip =geo_smk_pm_lag5_zip/10,
         geo_smk_pm10_lag6_zip =geo_smk_pm_lag6_zip/10,
         geo_smk_pm10_lag7_zip =geo_smk_pm_lag7_zip/10) %>%
  filter(!is.na(geo_smk_pm10_lag7_zip))

inhaler_ndc_lag_df$day[which(inhaler_ndc_lag_df$date=="2013-05-27")] <- "weekend"
inhaler_ndc_lag_df$day[which(inhaler_ndc_lag_df$date=="2013-07-04")] <- "weekend"
inhaler_ndc_lag_df$day[which(inhaler_ndc_lag_df$date=="2013-09-02")] <- "weekend"

which(colnames(inhaler_ndc_lag_df)=="geo_smk_pm10_zip") # 159
which(colnames(inhaler_ndc_lag_df)=="geo_smk_pm10_lag1_zip") # 160

inhaler_ndc_lag <- inhaler_ndc_lag_df %>%
  select(159, 160:166) # GWR-method

lag_max <- 7
n <- length(inhaler_ndc_lag[,1])

inhaler_ndc_lag <- as.matrix(inhaler_ndc_lag)

B <- ns(0:lag_max, df =3, intercept=TRUE) # or 3?

# the nrow B should equal ncol(pm)
dim(B)
dim(inhaler_ndc_lag)

inhaler_ndc_lag_B <- inhaler_ndc_lag%*%B
fit <- clogit(outcome ~ inhaler_ndc_lag_B + wrf_temp_zip + strata(personkey), inhaler_ndc_lag_df)

coef(fit)
summary(fit)
